Estimating duration in the presence of misclassification

### 2024-03-08 introduction.tex
Updates from Overleaf

COVID-19 (coronavirus disease 2019), the disease caused by the virus SARS-CoV-2 (severe acute respiratory syndrome coronavirus 2), is estimated to have killed approximately 15 million in the acutest phase (to the end of 2021) and 30 million in total~\autocite{whoCOVIDExcess,economistCOVIDExcess}.
History suggests severe pandemics are not rare (see \cref{intro:fig:pandemic-timeline}).
\begin{figure}
    \includegraphics[width=\textwidth]{introduction/pandemic-timeline}
    \caption[Timeline of pandemics.]{%
        Timeline of pandemics.
        See \url{https://ourworldindata.org/historical-pandemics} for details and sources.
        Licensed under \href{https://creativecommons.org/licenses/by/4.0/}{CC-BY} by Saloni Dattani, Klara Auerbach, Marwa Boukarim, and Max Roser.
    }
    \label{intro:fig:pandemic-timeline}
\end{figure}

Fully quantifying the burden of historical pandemics is challenging due to the absence of systematic records.
However, our best records suggests that, since 1600, they have (on average) occurred every four years with a mean death toll of 5.6 million~\autocite{maraniNovelEpidemics}.
Recent estimates suggest that the current mean impact of pandemics is over 2.5 million deaths per year~\autocite{madhavPandemicMortality}.

@@

Compartmental models can be fit to almost any data source, and commonly incorporate multiple data sources, known as \emph{data synthesis}, a technique reviewed by \textcite{birrellEvidence}.

\subsubsection{Application to SARS-CoV-2}

Many mechanistic modelling studies were conducted during the pandemic, with at least 100 articles published in the first six months of 2020 alone~\autocite{shankarSystematic}.
In the UK, SPI-M-O (Scientific Pandemic Influenza Group on Modelling, Operational sub-group) was a sub-committee of SAGE (the Scientific Advisory Group for Emergencies), generating evidence based on modelling for policy development~\autocite{medleySPIM,govSPIMO}.
For example, the stages of lifting the lockdown in 2021 each had a set of modelling studies informing the government's decision~\autocite{sageEvidence}.

Each week, the members of SPI-M would produce a set of estimates of the current state of the epidemic (\eg the value of $\Re$), known as \emph{nowcasting}.
The nowcasting estimates were based on mechanistic models.
These estimates, once any estimates unduly affected by issues (\eg data quality issues), were statistically combined into a consensus value~\autocite{parkCombining}.
The consensus value was published on \url{gov.uk} and informed the government's response to the epidemic~\autocite{govRnumber}.
The academic groups contributing estimates used a variety of models, reviewed by \textcite{royalSocietyRnumber}; see appendix 2 of that review for details of the specific implementations.
Most compartmental models used by SPI-M-O were age-stratified, as differences between age groups are important for both the transmission and severity of SARS-CoV-2 (see \cref{E-SEIR:sec:structured-populations}).

>>>>

COVID-19 (coronavirus disease 2019), the disease caused by the virus SARS-CoV-2 (severe acute respiratory syndrome coronavirus 2), is estimated to have killed approximately 15 million in the acutest phase (to the end of 2021) and 30 million in total~\autocite{whoCOVIDExcess,economistCOVIDExcess}.
History suggests severe pandemics are not rare (see \cref{intro:fig:pandemic-timeline}).
\begin{figure}
    \includegraphics[width=\textwidth]{introduction/pandemic-timeline}
    \caption[Timeline of pandemics.]{%
        Timeline of pandemics.
        See \url{https://ourworldindata.org/historical-pandemics} for details and sources.
        Licensed under \href{https://creativecommons.org/licenses/by/4.0/}{CC-BY} by Saloni Dattani, Klara Auerbach, Marwa Boukarim, and Max Roser.
    }
    \label{intro:fig:pandemic-timeline}
\end{figure}
Fully quantifying the burden of historical pandemics is challenging due to the absence of systematic records.
However, our best records suggests that, since 1600, they have (on average) occurred every four years with a mean death toll of 5.6 million~\autocite{maraniNovelEpidemics}.
Recent estimates suggest that the current mean impact of pandemics is over 2.5 million deaths per year~\autocite{madhavPandemicMortality}.

@@

Compartmental models can be fitted to almost any data source, and commonly incorporate multiple data sources, known as \emph{data synthesis}, a technique reviewed by \textcite{birrellEvidence}.

\subsubsection{Application to SARS-CoV-2}
{\color{red} you are missing a big opportunity here. You could elaborate a bit more the SPIMO part to explain what kind of data people used and introduce the question of whether the same evidence could be provided just by using data from a CIS type model. This is what motivates your research. Other people have used (only?) prevalence data (you give a lot of references below)..but have they recosntructed incidence and transmission only using prevalence data?}

Many studies involving mechanistic models were conducted during the pandemic, with at least 100 articles published in the first six months of 2020 alone~\autocite{shankarSystematic}.
In the UK, the Scientific Pandemic Influenza Group on Modelling, Operational sub-group (SPI-M-O) was a sub-committee of the Scientific Advisory Group for Emergencies (SAGE), using this type of models to generate evidence for policy development~\autocite{medleySPIM,govSPIMO}.
For example, each of the stages of lifting the lockdown in 2021 was supported by  modelling studies informing the government's decision~\autocite{sageEvidence}.

Each week, the members of SPI-M would produce a set of estimates of the current state of the epidemic (\eg the value of $\Re$), known as \emph{nowcasting}.
The nowcasting estimates were mainly based on mechanistic models.
These estimates, once any estimates unduly affected by issues (\eg data quality issues), were statistically combined into a consensus value~\autocite{parkCombining}.
The consensus value was published on \url{gov.uk} and informed the government's response to the epidemic~\autocite{govRnumber}.
The academic groups contributing estimates used a variety of models, reviewed by \textcite{royalSocietyRnumber} (see appendix 2 of that review for details of the specific implementations).
Most compartmental models used by SPI-M-O were age-stratified, as differences between age groups are important for both the transmission and severity of SARS-CoV-2 (see \cref{E-SEIR:sec:structured-populations}).
